Rubin Observatory Pointing Algorithm and Simulations

### 2025-01-30 ms.tex
general writing

%% If you want to create your own macros, you can do so
%% using \newcommand. Your macros should appear before
%% the \begin{document} command.
%%
\newcommand{\vdag}{(v)^\dagger}
\newcommand\aastex{AAS\TeX}
\newcommand\latex{La\TeX}

@@

xxx-Cite Eric's paper and LCOGT on integer programming.

Lots of buzz around AI models, but they are overblown \citep{Hicks24}

@@

\subsection{Deep Drilling Fields}
The LSST is planned to include several deep drilling fields. xxx--each DDF, maybe a cite for other studies of them. We plan to observe the Euclid Deep Field South (EDFS) as a pair of pointings.

@@

\begin{figure}
    \plotone{plots/hp2p_n1.pdf}
    \plotone{plots/hp2p_n5.pdf}
    \plotone{plots/nover_n1.pdf}
    \plotone{plots/nover_n5.pdf}
\end{figure}

@@

Describe the basic decision tree that gets built.

XXX--lay out what at baseline simulation looks like

\citet{Awan2016} for discussion of dithering strategies.

@@

\software{Astropy \citep{astropy2013, astropy2018}, 
          Healpy and HEALpix\footnote{\url{http://healpix.sourceforge.net}} \citep{healpix2005, Zonca2019},
          Numpy \citep{harris2020array},
          Matplotlib \citep{Hunter:2007},
          Scipy \citep{2020SciPy-NMeth},
          Pandas \citep{reback2022},
          Scikit-learn \citep{scikit-learn},
          OpenOrb \citep{Granvik2009},
          GNU Parallel \citep{Tange2023},
          rubin\_scheduler \citep{rubin_scheduler2024},
          rubin\_sim \citep{rubin_sim2024}}

@@

\begin{acknowledgments}
This work was facilitated through the use of advanced computational, storage, and networking infrastructure provided by the Hyak supercomputer system at the University of Washington.
\end{acknowledgments}

>>>>

%% If you want to create your own macros, you can do so
%% using \newcommand. Your macros should appear before
%% the \begin{document} command.
%%
\newcommand{\vdag}{(v)^\dagger}
\newcommand\aastex{AAS\TeX}
\newcommand\latex{La\TeX}
\usepackage{booktabs}

@@

xxx-Cite Eric's paper \citep{Bellm2019} and LCOGT \citep{Lampoudi2015} on integer programming.

There have been many recent developments using Machine Learning algorithms and Large Language Model AIs that have led many to believe a trained AI could be used to schedule observations. Unfortunately, the scale of LSST is so much larger than previous surveys it's not clear there is a data set of adequate size that could be used to train and LSST scheduling AI. There is also growing skeptisism around the feasability of making AIs that do not "hallucinate" \citep{Hicks24}.

@@

\subsection{Pre-Scheduled Observations}
The LSST is planned to include several deep drilling fields. xxx--each DDF, maybe a cite for other studies of them. We plan to observe the Euclid Deep Field South (EDFS) as a pair of pointings.

@@

xxx--discuss the general basis functions that typically are used for the MDP surveys.

Figure~\ref{fig:tesselation} illustrates our nightly dithering scheme. \citet{Awan2016} for discussion of dithering strategies.

\begin{figure}
    \epsscale{0.5}
    \plottwo{plots/hp2p_n1.pdf}{plots/hp2p_n5.pdf} \\
    \plottwo{plots/nover_n1.pdf}{plots/nover_n5.pdf}
    \epsscale{1}
    \caption{Examples of how the sky tesselation is rotated on different nights. Figures show a gnomic projection of the south celectial pole for declinations below -60 degrees. In the top panels, we color HEALpixels based on which pointing they would be assigned to. Bottom panels show the overlaps between pointings. Night 1 of the survey is on the left, while night 5 is on the right. \label{fig:tesselation}}
\end{figure}

\subsection{Combined MDP and Prescheduled}

We can combine the above two techniques. In order to force observations on certain time scales, we can use the MDP to observe an area of the sky, and once those observations are completed, spawn new scheduled observations on the same area to be completed later.

\subsection{Target of Opportunity}

There are a number of potential Target of Opportunity

@@

\subsection{Scheduler Decision Tree}
Describe the basic decision tree that gets built. Tier 1 are ToO observations. Tier 2 are Roman overlap observations. Tier 3 are Deep Drilling Field observations. Tier 4 are observations dedicated to observing time scales of several hours. Tier 5 are taking paired observations with 33 minute time gaps. Tier 6 is similar, but for 15 minute gaps. Tier 7 is earth-interior solar system objects, so sets of 4 observations taken at high airmass in twilight time

\subsection{}

@@

\software{Astropy \citep{astropy2013, astropy2018}, \\
          Healpy and HEALpix\footnote{\url{http://healpix.sourceforge.net}} \citep{healpix2005, Zonca2019},
          Numpy \citep{harris2020array}, \\
          Matplotlib \citep{Hunter:2007},
          Scipy \citep{2020SciPy-NMeth},
          Pandas \citep{reback2022}, \\
          Scikit-learn \citep{scikit-learn},
          OpenOrb \citep{Granvik2009}, 
          GNU Parallel \citep{Tange2023}, \\
          rubin\_scheduler \citep{rubin_scheduler2024},
          rubin\_sim \citep{rubin_sim2024}}

@@

\begin{acknowledgments}
This work was facilitated through the use of advanced computational, storage, and networking infrastructure provided by the Hyak supercomputer system at the University of Washington.

This work used the resources of the SLAC Shared Science Data Facility (S3DF) at SLAC National Accelerator Laboratory. S3DF is a shared High-Performance Computing facility, operated by SLAC, that supports the scientific and data-intensive computing needs of all experimental facilities and programs of the SLAC National Accelerator Laboratory. SLAC is operated by Stanford University for the U.S. Department of Energy’s Office of Science.
\end{acknowledgments}
